---
title: "hw4"
author: "Jenna"
date: "10/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(gbm)
```

```{r include=FALSE}
hw4 = read.csv("source_data/hw4.csv")
analysis = hw4 %>% mutate(Male=(Gender=="Male"))
```
# Problem 1

Build a glm in R to classifier individuals as either Male or Female based on their weight and height.

What is the accuracy of the model?

```{r}
set.seed(1107)
train=sample(c(TRUE,FALSE),nrow(analysis),rep=TRUE)
test=(!train)

train = analysis[train,]
test = analysis[test,]

fit = glm(Male~Weight+Height,data=train,family=binomial())
prediction=(predict(fit,type="response",test)>0.5)
results=as.data.frame(cbind(Truth=test$Male,Prediction=prediction)) %>% mutate(correct=(Truth==Prediction))
table(results$correct)
131/(128+131)
```
# Problem 2

Use the 'gbm' package to train a similar model. Don't worry about hyper parameter tuning for now.

What is the accuracy of the model?

```{r}

```


# Problem 3

Filter the data set so that it contains only 50 Male examples. Create a new model for this data set. What is the F1 Score of the model?

```{r}
male = analysis %>% filter(Male==1)
female = analysis %>% filter(Male==0)
filter = rbind(male[1:50,],female)

set.seed(132)
train=sample(c(TRUE,FALSE),nrow(filter),rep=TRUE)
test=(!train)

train = analysis[train,]
test = analysis[test,]

fit2 = glm(Male~Weight+Height,data=train,family=binomial())
prediction2=(predict(fit2,type="response",test)>0.5)
results2=as.data.frame(cbind(Truth=test$Male,Prediction=prediction2)) %>% mutate(correct=(Truth==Prediction))
table(results2$correct)
117/(130+117)
```


# Problem 4

For the model in the previous example plot an ROC curve. What does this ROC curve mean?

# Problem 5

Using K-Means, cluster the same data set. Can you identify the clusters with the known labels? Provide an interpretation of this result.

